---
title: "Data Figures Script for the Appendix"
author: "Stephanie Copeland & Samantha Sambado"
date: "2024-08-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## upload neccessary packages
library(tidyverse)
library(readr)
library(ggplot2)
library(dplyr)
library(plyr)
library(knitr)
library(lme4)
library(jtools)
library(MuMIn)
library(rstatix)
#install.packages("glmmTMB", type="source")
library(glmmTMB)
library(pscl)
#install.packages("ggsignif")
library(ggsignif)
#install.packages("ggpubr")
library(ggpubr)
library(patchwork)

```

```{r}
## upload data
tick <- read_csv("Tejon_MeanComparisons_Dataset.csv")
dim(tick) # 1449   11

#plot = herbivore exclosure treatment
# Partial = wildlife only (W)
# Open = wildlife & cattle (CW)
# Total = minus wildlife & cattle (-CW)

#site = climate level 
# Arid
# Intermediate 
# Mesic

## working dataframe
Tick <- tick

### format data
Tick$deoc <- as.integer(Tick$deoc)
Tick$site <- as.factor(Tick$site)
Tick$plot <- as.factor(Tick$plot)
Tick$year <- as.factor(Tick$year)
Tick$month <- as.factor(Tick$month)

## change any plots in 2016 to 'Open', fencing was established in November of 2016
Tick <- Tick %>%
  mutate(plot = if_else(year == "2016", "Open", plot))

# remove pre-treatment 2016 data for some of the stats
# summary analyses includes all tick species collected in drags
# effect and regression analyses include only D. occidentalis counts from 2017 onward
Tick1 <- Tick %>%
  filter(year != "2016")
dim(Tick1)

Tick1$deoc <- as.integer(Tick1$deoc)
Tick1$site <- as.factor(Tick1$site)
Tick1$plot <- as.factor(Tick1$plot)
Tick1$year <- as.factor(Tick1$year)
Tick1$month <- as.factor(Tick1$month)
```

## Appendix Figures ##

### Figure S1(a & b) ###

Mean observed vs. Mean predicted for climate and herbivory model fit

**Fig S1a**
```{r}
Tickmod <- na.omit(Tick1)

#best fit model
mod_2_4 <- glmmTMB(deoc ~ plot*site*year + (1|month) + (1|plotID),
                  family = nbinom2(), 
                  data = Tick1)

#make predictions
Tickmod$predicted <- predict(mod_2_4, type = "response")

# Get predictions and standard errors
predictions <- predict(mod_2_4, type = "response", se.fit = TRUE)
Tickmod$predicted <- predictions$fit
Tickmod$se <- predictions$se.fit

### format data
Tickmod$deoc <- as.integer(Tickmod$deoc)
Tickmod$site <- as.factor(Tickmod$site)
Tickmod$plot <- as.factor(Tickmod$plot)
Tickmod$year <- as.factor(Tickmod$year)
Tickmod$month <- as.factor(Tickmod$month)

#aggregate data
modagg_data <- Tickmod %>%
  dplyr::group_by(plot, site, year) %>%
  dplyr::summarize(mean_observed = mean(deoc, na.rm = TRUE),
            mean_predicted = mean(predicted, na.rm = TRUE),
            .groups = 'drop')

FigS1a <- ggplot(modagg_data, aes(x = mean_observed, y = mean_predicted, color = plot)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE, color = "grey", size = 0.5) +
  labs(x = "Mean Observed Values",
       y = "Mean Predicted Values") +
    scale_color_manual(values = c("Open" = "darkseagreen1", 
                                "Partial" = "darkseagreen3", 
                                "Total" = "darkseagreen4"),
                        labels = c("Open" = "CW", 
                                  "Partial" = "W", 
                                  "Total" = "-CW"),
                       breaks = c("Open", "Partial", "Total")) +
      annotate("text", x = -Inf, y = Inf, label = "a.", hjust = -0.8, vjust = 2.0, size = 4, family = "Times New Roman") +
  guides(color = guide_legend(title = "Herbivore Treatment")) + 
  theme_bw()+
  theme(
    legend.position =  'bottom',
    text = element_text(family = "Times New Roman"), 
    axis.text.x = element_text(size = 10),          
    axis.text.y = element_text(size = 10),           
    axis.title.x = element_text(size = 12),          
    axis.title.y = element_text(size = 12),          
    plot.title = element_blank()                     
  )
FigS1a
```

**Fig S1b**
```{r}
FigS1b <- ggplot(modagg_data, aes(x = mean_observed, y = mean_predicted, color = site)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE, color = "grey", size = 0.5) +
  labs(x = NULL,
       y = NULL) +
    scale_color_manual(values = c("Arid" = "brown1", 
                                "Intermediate" = "tan1", 
                                "Mesic" = "lightgoldenrod1"),
                       breaks = c("Arid", "Intermediate", "Mesic")) +
    annotate("text", x = -Inf, y = Inf, label = "b.", hjust = -0.8, vjust = 2.0, size = 4, family = "Times New Roman") +
  guides(color = guide_legend(title = "Climate Level")) + 
  theme_bw()+
  theme(
    legend.position =  'bottom',
    text = element_text(family = "Times New Roman"), 
    axis.text.x = element_text(size = 10),           
    axis.text.y = element_text(size = 10),          
    axis.title.x = element_text(size = 12),          
    axis.title.y = element_text(size = 12),         
    plot.title = element_blank()                     
  )
FigS1b
```

combine plots
```{r}
# Combine the two plots side by side
combined_plot <- FigS1a + FigS1b + plot_layout(ncol = 2)

# Display the combined plot
combined_plot

#ggsave("FigS1a_b.png", combined_plot, width = 7, dpi = 600)
```

### Figure S2 ###

Monthly log of tick abundance 

```{r}
Tickmonth <- Tick1

Tickmonth$month <- factor(Tickmonth$month, 
                        levels = c('January', 'February', 'March',
                                   'April', 'May', 'June', 'July',
                                   'August', 'September', 'October',
                                   'November', 'December'))

Tickmonth$site <- factor(Tickmonth$site, 
                    levels = c('Arid', 'Intermediate', 'Mesic'))


Tickmonth$plot <- recode(Tickmonth$plot,
                        Open = "CW",
                        Partial = "W",
                        Total = "-CW")

Tickmonth$plot <- factor(Tickmonth$plot,
                         levels = c("CW", "W", "-CW"))

FigS2 <- ggplot(Tickmonth, aes(x = site, y = log(total +1), fill = plot))+
  facet_wrap(~ month, ncol = 6)+
  geom_boxplot()+
  theme_bw() +
  scale_x_discrete(labels = c("A", "I", "M")) +
  theme(legend.position =  'bottom',
    text = element_text(family = "Times New Roman"), 
    axis.text.x = element_text(size = 10),           
    axis.text.y = element_text(size = 10),          
    axis.title.x = element_text(size = 12),          
    axis.title.y = element_text(size = 12),         
    plot.title = element_text(size = 12))+
  labs(x = 'Climate Level', y = 'Natural Log of Relative Tick Abundance (2017 - 2019)', fill = "Herbivore Treatment")+
  scale_fill_manual(values=c('darkseagreen1','darkseagreen3','darkseagreen4'))

FigS2

#ggsave("FigS2.png", FigS2, width = 7, dpi = 600)
```

**Fig S3 (a & b): Lizard population estimates by climate level and herbivore treatment**

```{r}
## read in data
pop <- read_csv("Tejon_MixedModels_Dataset.csv")

pop1 <- pop %>%
  filter(year != "2016")

pop1$site <- as.factor(pop1$site)
pop1$plot <- as.factor(pop1$plot)
```

```{r}
#remove NAs
pop1liz <- pop1 %>%
  filter(year != "2019") #remove 2019

pop1liz <- pop1liz[!is.na(pop1liz$liz_estimate), ]

pop1liz$liz_estimate <- as.numeric(pop1liz$liz_estimate)

summary(pop1liz$liz_estimate)

```

```{r}
FigS3a <- ggplot(pop1liz, aes(x = site, y = liz_estimate, fill = site))+
  geom_boxplot()+
  geom_point(alpha = .15)+
  #geom_signif(comparisons = list(c("Arid", "Intermediate")),
  #map_signif_level=TRUE)+
  theme_bw() +
    theme(legend.position =  "none",
    text = element_text(family = "Times New Roman"), 
    axis.text.x = element_text(size = 10),           
    axis.text.y = element_text(size = 10),          
    axis.title.x = element_text(size = 12),          
    axis.title.y = element_text(size = 12),         
    plot.title = element_blank())+
  scale_fill_manual(values=c('brown1','tan1','lightgoldenrod1'))+
  labs(x = "Climate Level", y = "Lizard Population Estimates")+
  annotate("text", x = -Inf, y = Inf, label = "a.", hjust = -0.8, vjust = 1.5, size = 4, family = "Times New Roman")

FigS3b <- ggplot(pop1liz, aes(x = plot, y = liz_estimate, fill = plot))+
  geom_boxplot()+
  geom_point(alpha = .15)+
  #geom_signif(comparisons = list(c("Partial", "Total")),
  #map_signif_level=TRUE)+
  theme_bw() +
  scale_x_discrete(labels = c("CW", "W", "-CW")) +
  theme(legend.position =  "none",
    text = element_text(family = "Times New Roman"), 
    axis.text.x = element_text(size = 10),           
    axis.text.y = element_text(size = 10),          
    axis.title.x = element_text(size = 12),          
    axis.title.y = element_text(size = 12),         
    plot.title = element_blank())+
  labs(x = "Herbivore Treatment", y = NULL)+
  scale_fill_manual(values=c('darkseagreen1','darkseagreen3','darkseagreen4'))+
  annotate("text", x = -Inf, y = Inf, label = "b.", hjust = -0.8, vjust = 1.5, size = 4, family = "Times New Roman")

```
combine plots
```{r}
# Combine the two plots side by side
combined_plot <- FigS3a + FigS3b + plot_layout(ncol = 2)

# Display the combined plot
combined_plot

#ggsave("FigS3a_b.png", combined_plot, width = 7, dpi = 600)
```

### Figure S4 ###

**Figure S4 (a & b): Small Mammal Population Estimates by climate level and herbivore treatment**

```{r}
#remove NAs
pop1mam <- pop1 %>%
  filter(year != "2019") #removed 2019

pop1mam <- pop1mam[!is.na(pop1mam$mam_estimates), ]

pop1mam$mam_estimates <- as.numeric(pop1mam$mam_estimates)
```

```{r}
FigS4a <- ggplot(pop1mam, aes(x = site, y = mam_estimates, fill = site))+
  geom_boxplot()+
  geom_point(alpha = .15)+
  #geom_signif(comparisons = list(c("Arid", "Intermediate"), c("Intermediate", "Mesic"), c("Arid", "Mesic")),
  #map_signif_level=TRUE)+
  theme_bw() +
    theme(legend.position =  "none",
    text = element_text(family = "Times New Roman"), 
    axis.text.x = element_text(size = 10),           
    axis.text.y = element_text(size = 10),          
    axis.title.x = element_text(size = 12),          
    axis.title.y = element_text(size = 12),         
    plot.title = element_blank())+
  scale_fill_manual(values=c('brown1','tan1','lightgoldenrod1'))+
  labs(x = "Climate Level", y = "Mammal Population Estimates")+
  annotate("text", x = -Inf, y = Inf, label = "a.", hjust = -0.8, vjust = 1.5, size = 4, family = "Times New Roman")


FigS4b <- ggplot(pop1mam, aes(x = plot, y = mam_estimates, fill = plot))+
  geom_boxplot()+
  geom_point(alpha = .15)+
  #geom_signif(comparisons = list(c("Open", "Partial"), c("Partial", "Total"), c("Open", "Total")),
  #map_signif_level=TRUE)+
  theme_bw() +
  scale_x_discrete(labels = c("CW", "W", "-CW")) +
    theme(legend.position =  "none",
    text = element_text(family = "Times New Roman"), 
    axis.text.x = element_text(size = 10),           
    axis.text.y = element_text(size = 10),          
    axis.title.x = element_text(size = 12),          
    axis.title.y = element_text(size = 12),         
    plot.title = element_blank())+
  scale_fill_manual(values=c('darkseagreen1', 'darkseagreen3', 'darkseagreen4'))+
  labs(x = "Herbivore Treatment", y = NULL)+
  annotate("text", x = -Inf, y = Inf, label = "b.", hjust = -0.8, vjust = 1.5, size = 4, family = "Times New Roman")

```
combine plots
```{r}
# Combine the two plots side by side
combined_plot <- FigS4a + FigS4b + plot_layout(ncol = 2)

# Display the combined plot
combined_plot

#ggsave("FigS4a_b.png", combined_plot, width = 7, dpi = 600)
```