---
title: "Data Cleaning"
author: "Stephanie Copeland & Samantha Sambado"
date: "2021-10-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readr)
library(dplyr)
library(plyr)
library(here)
library(Rcapture)
```

## Dataset 1 -- Tick Density

Final version of exported datafile = "Tejon_MeanComparisons_Dataset.xlsx"

###n Step 1. Merge Data \###

```{r}
threeyrs <- read_csv("Total_Ticks_2016_to_2018.csv")
oneyr <- read_csv("Tick_19.csv")
```

```{r}
# get rid of survery column in threeyr since it wasn't completed in oneyrs 
threeyrs <- threeyrs %>% 
  select(-Survey)
```

```{r}
# make sure names are the same
## 2016 - 2018
names(threeyrs) <- tolower(names(threeyrs))

# change column names
names(threeyrs)[6] <- "total"
names(threeyrs)[7] <- "deoc"
names(threeyrs)[8] <- "ipac"
names(threeyrs)[9] <- "deva"
names(threeyrs)[10] <- "other"
```

```{r}
## 2019
# make sure names are the same
names(oneyr) <- tolower(names(oneyr))

# change column names
names(oneyr)[6] <- "total"
names(oneyr)[7] <- "deoc"
names(oneyr)[8] <- "ipac"
names(oneyr)[9] <- "deva"
names(oneyr)[10] <- "other"
```

```{r}
tick <- rbind(threeyrs, oneyr)
```

```{r}
## add plotID

tick <- transform(tick, plotID = as.numeric(interaction(site, plot,block, drop = TRUE)))
```

### Step 2. Check Data Structures

```{r}
str(tick) 

tick$year <- as.numeric(tick$year)
tick$site <- as.factor(tick$site)
tick$plot <- as.factor(tick$plot)

## Changed to Tick from tick due to it being called that way in some analyses

Tick <- tick

#install.packages("writexl")
library(writexl)

#write_xlsx(Tick, path = "Tejon_MeanComparisons_Dataset.xlsx",
          #col_names = TRUE, format_headers = FALSE)
```

## Dataset 2 -- Mammal & Lizard Data

Final version of exported file = "Tejon_MixedModels_Dataset.xlsx"

### Population Estimates for Mammals

```{r}
mamcounts <- read_csv("mammal_counts_02182021.csv")
```

```{r}
mamcounts$day1 <- as.numeric(mamcounts$day1)
mamcounts$day2 <- as.numeric(mamcounts$day2)
mamcounts$day3 <- as.numeric(mamcounts$day3)
mamcounts$site <- as.factor(mamcounts$site)
mamcounts$plot <- as.factor(mamcounts$plot)
mamcounts$block <- as.numeric(mamcounts$block)
```

#### Arid

```{r}
library(Rcapture)

AridO <- mamcounts %>%
  filter(site == "Arid", plot =="Open", block == 1) %>%
  select(day1, day2, day3)
  
abundance0 = closedp.t(AridO, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundance0$results)
models <- models[order(models$BIC), ]

models[1,1] #select total abundance

## Open (Wildlife and Cattle)
#Block 1 = 2
#Block 2 = 5
#Block 3 = 15

AridP <- mamcounts %>%
  filter(site == "Arid", plot =="Partial", block == 1) %>%
  select(day1, day2, day3)
  
abundance0 = closedp.t(AridP, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundance0$results)
models <- models[order(models$BIC), ]

models[1,1] #select total abundance

## Partial (Wildlife)
# Block 1 = 16
# Block 2 = 46, lowest BIC was negative so chose next lowest model M0
# Block 3 = Estimates were in the 500000000 so we chose to omit this block 

AridT <- mamcounts %>%
  filter(site == "Arid", plot =="Total", block == 1) %>%
  select(day1, day2, day3)
  
abundance0 = closedp.t(AridT, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundance0$results)
models <- models[order(models$BIC), ]

models[1,1] #select total abundance

## Total (-Cattle & Wildlife)
# Block 1 = 16
# Block 2 = 41
# Block 3 = 16

```

#### Intermediate

```{r}
IntermediateO <- mamcounts %>%
  filter(site == "Intermediate", plot =="Open", block == 2) %>%
  select(day1, day2, day3)
  
abundanceO = closedp.t(IntermediateO, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceO$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

## Open (Wildlife and Cattle)
# Block 1 = no mammals trapped at block 1, Intermediate, C&W
# Block 2 = 6
# Block 3 = estimates were in the 7e09 so we chose to omit this block 

IntermediateP <- mamcounts %>%
  filter(site == "Intermediate", plot =="Partial", block == 1) %>%
  select(day1, day2, day3)
  
abundanceO = closedp.t(IntermediateP, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceO$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

## Partial (Wildlife)
# Block 1 = 8
# Block 2 = 2
# Block 3 = estimates were in 700 so omitted this data

IntermediateT <- mamcounts %>%
  filter(site == "Intermediate", plot =="Total", block == 1) %>%
  select(day1, day2, day3)
  
abundanceO = closedp.t(IntermediateT, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceO$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

## Total (-Wildlife and Cattle)
# Block 1 = 32
# Block 2 = estimates over 5 mil so omitted
# Block 3 = 2
```

#### Mesic

```{r}
MesicO <- mamcounts %>%
  filter(site == "Mesic", plot =="Open", block == 1) %>%
  select(day1, day2, day3)
  
abundanceO = closedp.t(MesicO, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceO$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance

## Open (Wildlife and Cattle)
# Block 1 = 29
# Block 2 = 2
# Block 3 = 13

MesicP <- mamcounts %>%
  filter(site == "Mesic", plot =="Partial", block == 1) %>%
  select(day1, day2, day3)
  
abundanceO = closedp.t(MesicP, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceO$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance

## Partial (Wildlife)
# Block 1 = 216
# Block 2 = 16
# Block 3 = estimates were 10mil so omitted them

MesicT <- mamcounts %>%
  filter(site == "Mesic", plot =="Total", block == 1) %>%
  select(day1, day2, day3)
  
abundanceO = closedp.t(MesicT, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceO$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance

## Total (-Wildlife and Cattle)
# Block 1 = 63
# Block 2 = 8
# Block 3 = 18

```

### Population Estimates for Lizards

```{r}
lizcounts <- read_csv("lizardcounts_02172021.csv")
```

```{r}
names(lizcounts) <- tolower(names(lizcounts))

lizcounts$day1 <- as.numeric(lizcounts$day1)
lizcounts$day2 <- as.numeric(lizcounts$day2)
lizcounts$day3 <- as.numeric(lizcounts$day3)
lizcounts$site <- as.factor(lizcounts$site)
lizcounts$plot <- as.factor(lizcounts$plot)
lizcounts$block <- as.numeric(lizcounts$block)
```

```{r}
AridO <- lizcounts %>%
  filter(site == "Arid", plot =="Open", block == 1) %>%
  select(day1, day2, day3)
  
abundance0 = closedp.t(AridO, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundance0$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

## Open (Wildlife and Cattle)
# Block 1 = 52
# Block 2 = 61
# Block 3 = 83

AridP <- lizcounts %>%
  filter(site == "Arid", plot =="Partial", block == 1) %>%
  select(day1, day2, day3)
  
abundance0 = closedp.t(AridP, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundance0$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

## Partial (Wildlife)
# Block 1 = 74
# Block 2 = 73
# Block 3 = 70

AridT <- lizcounts %>%
  filter(site == "Arid", plot =="Total", block == 1) %>%
  select(day1, day2, day3)
  
abundance0 = closedp.t(AridT, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundance0$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

## Total (-Wildlife and Cattle)
# Block 1 = 68
# Block 2 = 61
# Block 3 = 89

```

#### Intermediate

```{r}
IntermediateO <- lizcounts %>%
  filter(site == "Intermediate", plot =="Open", block == 1) %>%
  select(day1, day2, day3)
  
abundanceO = closedp.t(IntermediateO, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceO$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance 

## Open (Wildlife and Cattle)
# Block 1 = 130
# Block 2 = 85
# Block 3 = 208

IntermediateP <- lizcounts %>%
  filter(site == "Intermediate", plot =="Partial", block == 2) %>%
  select(day1, day2, day3)
  
abundanceO = closedp.t(IntermediateP, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceO$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance

## Partial (Wildlife)
# Block 1 = 134
# Block 2 = 48
# Block 3 = 163

IntermediateT <- lizcounts %>%
  filter(site == "Intermediate", plot =="Total", block == 1) %>%
  select(day1, day2, day3)
  
abundanceO = closedp.t(IntermediateT, dfreq = FALSE, dtype="hist") 
models <- as.data.frame(abundanceO$results)
models <- models[order(models$BIC), ]

models[1,1] # select total abundance

## Total (-Wildlife and Cattle)
# Block 1 = 138, lowest BIC was negative so chose next lowest model Mb
# Block 2 = 85
# Block 3 = 177
```

Did not sample lizards at the Mesic climate location.

**All of these result values were entered into an Excel spreadsheet titled:** `tejon_mamliz_popestimates.xlsx`

## Add lizard and mammal estimates with tick counts.

```{r}
# log transform tick count data
Tick$log_total <- log(Tick$total + 1)

# vertebrate population estimates
pop_estimates <-read_csv("tejon_mamliz_popestimates.csv")

# add pop estimates to tick data
vert.tick <- join(Tick, pop_estimates)
```

### Tidy Data

```{r}
vert.tick$year <- as.factor(vert.tick$year)
vert.tick$block <- as.factor(vert.tick$block)
vert.tick$plotID <- as.factor(vert.tick$plotID)

vert.tick$month <- factor(vert.tick$month, 
                        levels = c('January', 'February', 'March',
                                   'April', 'May', 'June', 'July',
                                   'August', 'September', 'October',
                                   'November', 'December'))
vert.tick$site <- factor(vert.tick$site, 
                    levels = c('Arid', 'Intermediate', 'Mesic'))

vert.tick$plot <- factor(vert.tick$plot, 
                    levels = c('Open', 'Partial', 'Total'))

#write_xlsx(vert.tick, path = "Tejon_MixedModels_Dataset.xlsx",
           #col_names = TRUE, format_headers = FALSE)
```
